---
title: "Clean tidycensus maps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#packages needed
```{r}
library(tidycensus)
#library(tidyverse)
library(tmap)
library(tmaptools)
library(sf)
library(ggplot2)
library(dplyr)
library(mapview)
library(tigris)
library(leaflet)
```

###Race###

map for race in Hampden county by tract
```{r}
#getting all the variables for race from census
racevars <- c(White = "B02001_002", 
              Black = "B02001_003", 
              Asian = "B02001_005", 
              Hispanic = "B03002_001",
              Native_american = "B02001_004",
              Native_hawaian = "B02001_006",
              someOtherRace = "B02001_007")

race_tract <- get_acs(geography = "tract", variables = racevars, 
                  state = "MA", county = "Hampden County", geometry = TRUE,
                  summary_var = "B01003_001")  #what is the summary variable again

county_race_map <- race_tract %>%
                      mutate(pct = 100 * (estimate /summary_est)) %>% #moe in the df is margin of error
                      ggplot(aes(fill = pct)) + #is a continuous scale so can't have discrete values
                      facet_wrap(~variable) +
                      geom_sf(color = NA) +
                      coord_sf(crs = 26918) + 
                      scale_fill_viridis_c()

class(county_race_map)

county_race_map

```

map for race by place- problem is that place is available by state, not county so we get map of the entire state- not helpful
```{r}
race_place <- get_acs(geography = "place", variables = racevars, 
                  state = "MA", geometry = TRUE,
                  summary_var = "B01003_001")  #what is the summary variable again

race_place %>%
  mutate(pct = 100 * (estimate /summary_est)) %>% #moe in the df is margin of error
  ggplot(aes(fill = pct)) + #is a continuous scale so can't have discrete values
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  coord_sf(crs = 26918) + 
  scale_fill_viridis_c()

```

race maps with cities more clear
```{r}
race_block <- get_acs(geography = "block group", variables = racevars, 
                  state = "MA", county = "Hampden County",geometry = TRUE,
                  summary_var = "B01003_001")  #summary variable is total pop

options(tigris_use_cache = TRUE)

race_block %>%
  mutate(pct = 100 * (estimate /summary_est)) %>% #moe in the df is margin of error
  ggplot(aes(fill = pct)) + #is a continuous scale so can't have discrete values
  facet_wrap(~variable) +
  geom_sf(color = NA) +
  coord_sf(crs = 26918) + 
  scale_fill_viridis_c()
```

***Individual race maps***

--Largest racial populations--
(consider adding asian)

-Springfield: white, black, hispanic

-Chicopee: white, black, hispanic

-Holyoke:white, black, hispanic


```{r}
###white
white <- get_acs(geography = "place", #census tract-works for individual cities
                    variables = "B02001_002", #only white
                    state = "MA",
                    #county = "Hampden County",
                    year = 2018, #is also the defualt
                    geometry = TRUE,
                    survey = "acs5",
                    summary_var = "B01003_001") 

white <- white %>%
  mutate(pct = 100*(estimate/summary_est))



###black

black <- get_acs(geography = "place", #census tract-works for individual cities
                    variables = "B02001_003", #only african american
                    state = "MA",
                    #county = "Hampden County",
                    year = 2018,
                    geometry = TRUE,
                    survey = "acs5",
                    summary_var = "B01003_001") 

black <- black %>%
  mutate(pct = 100*(estimate/summary_est))



###hispanic

hispanic <- get_acs(geography = "place", #census tract-works for individual cities
                    variables = "B03002_001", #hispanic or latino origin by race
                    state = "MA",
                    #county = "Hampden County",
                    year = 2018,
                    geometry = TRUE,
                    survey = "acs5",
                    summary_var = "B01003_001") 

hispanic <- hispanic %>%
  mutate(pct = 100*(estimate/summary_est)) #pct is 100. how can summary and hispanic estimates be the same???

View(black)
View(hispanic)

#map output
mapview(white, zcol = "pct")
mapview(black, zcol = "pct")
mapview(hispanic, zcol = "pct")



```


###Income###

Income for the cities

```{r}
#people with no income
no_income <- get_acs(geography = "place", #census tract-works for individual cities
                    variables = "B06010_002", 
                    state = "MA",
                    #county = "Hampden County",
                    year = 2018, #is also the defualt
                    geometry = TRUE,
                    survey = "acs5",
                    summary_var = "B06010_001")

no_income <- no_income %>%
  mutate(pct = 100*(estimate/summary_est))

#$1 to $9,999 or loss
low_income1 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_004", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

low_income1 <- low_income1 %>%
  mutate(pct = 100*(estimate/summary_est))

#$10,000 to $14,999
low_income2 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_005", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

low_income2 <- low_income2 %>%
  mutate(pct = 100*(estimate/summary_est))

#$15,000 to $24,999
low_income3 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_006", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

low_income3 <- low_income3 %>%
  mutate(pct = 100*(estimate/summary_est))

#$25,000 to $34,999
mid_income1 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_007", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

mid_income1 <- mid_income1 %>%
  mutate(pct = 100*(estimate/summary_est))

#$35,000 to $49,999
mid_income2 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_008", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

mid_income2 <- mid_income2 %>%
  mutate(pct = 100*(estimate/summary_est))

#$50,000 to $64,999
mid_income3 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_009", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

mid_income3 <- mid_income3 %>%
  mutate(pct = 100*(estimate/summary_est))

#$65,000 to $74,999
high_income1 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_010",
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

high_income1 <- high_income1 %>%
  mutate(pct = 100*(estimate/summary_est))

#$75,000 or more
high_income2 <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B06010_011", 
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B06010_001") 

high_income2 <- high_income2 %>%
  mutate(pct = 100*(estimate/summary_est))


mapview(no_income, zcol = "pct")
mapview(low_income1, zcol = "pct")
mapview(low_income2, zcol = "pct")
mapview(low_income3, zcol = "pct")
mapview(mid_income1, zcol = "pct")
mapview(mid_income2, zcol = "pct")
mapview(mid_income3, zcol = "pct")
mapview(high_income1, zcol = "pct")
mapview(high_income2, zcol = "pct")


```




Income by race


##Health care##

-don't understand the health care but what we can do is show health care amenities (hospitals in each city)

##Education level##

```{r}
ged_or_similar <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_018", #GED or alternative
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

ged_or_similar <- ged_or_similar %>%
  mutate(pct = 100*(estimate/summary_est))

regHSdiploma <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_017", #regular high school diploma
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

regHSdiploma <- regHSdiploma %>%
  mutate(pct = 100*(estimate/summary_est))

associate <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_021", #associate's degres
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

associate <- associate %>%
  mutate(pct = 100*(estimate/summary_est))

bachelor <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_022", #bachelor's
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

bachelor <- bachelor %>%
  mutate(pct = 100*(estimate/summary_est))

masters <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_023", #master's
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

masters <- masters %>%
  mutate(pct = 100*(estimate/summary_est))

professional_school <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_024", #professional school degree
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

professional_school <- professional_school %>%
                mutate(pct = 100*(estimate/summary_est))

doctorate <- get_acs(geography = "place", #census tract-works for individual cities
                      variables = "B15003_025", #doctorate degree
                      state = "MA",
                      #county = "Hampden County",
                      year = 2018, #is also the defualt
                      geometry = TRUE,
                      survey = "acs5",
                      summary_var = "B15003_001") #educational attainment for the population 25 years and over

doctorate <- doctorate %>%
  mutate(pct = 100*(estimate/summary_est))

mapview(ged_or_similar, zcol = "pct")
mapview(regHSdiploma, zcol = "pct")
mapview(associate, zcol = "pct")
mapview(bachelor, zcol = "pct")
mapview(masters, zcol = "pct")
mapview(professional_school, zcol = "pct")
mapview(doctorate, zcol = "pct")

```


